---
title: "Practice Exam"
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_depth: 1
    fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this practice exam, we mainly focus on data cleaning and data visualizations.

For problems 2 and 3, you should mainly use **dplyr** and **ggplot2** to construct some plots and also provide **brief interpretations** about your findings.

## Problem 1: Dataset Import & Cleaning

The data comes from a global company, including orders from 2012 to 2015. Import the dataset **Order** and do some basic EDA. 

Check **"Profit"** and **"Sales"** in the dataset, convert these two columns to numeric data.

```{r prob1}
library(data.table)
library(VIM)
library(caret)
library(ggplot2)
library(dplyr)

orders = fread('./data/Orders.csv') # stringsAsFactors set to False already
#View(orders)
head(orders)
#tail(orders)
sapply(orders, class)
names(orders)
str(orders)
dim(orders)   #nrow() & ncol() show same
summary(orders)

#order.date and ship.date should be datetime
orders$Ship.Date = as.Date(orders$Ship.Date, '%m/%d/%y')
orders$Order.Date = as.Date(orders$Order.Date, '%m/%d/%y')

#profit and sales should be numeric
orders$Profit <- as.numeric(sub("\\$","", orders$Profit))
orders$Sales <- as.numeric(sub("\\$","", orders$Sales))

sapply(orders, class)  #checks out 

#histograms of numeric
hist(orders$Discount)
hist(orders$Profit)    # some neg profit ??
hist(orders$Sales)    #long tail
hist(orders$Shipping.Cost)


#distinct categories
unique(orders$Category)
unique(orders$Sub.Category)
unique(orders$Market)
unique(orders$Region)
#boxplots of characters
boxplot(orders$Profit ~ orders$Category, col="purple", ylim=c(-200,400))
boxplot(orders$Profit ~ orders$Sub.Category, col="blue", ylim=c(-200,400))
boxplot(orders$Profit ~ orders$Region, col="red", ylim=c(-200,400))
boxplot(orders$Profit ~ orders$Market, col="green", ylim=c(-200,400))
        
        
#checking missingness
sapply(orders, function(x) sum(is.na(x)))
#postal.code missing 41296 (80%)
#sales missing 2626 (5%)
#profit missing 225 (0.4%)
round(sapply(orders, function(x) sum(is.na(x)))/dim(orders)[1]*100, 1)
sum(!complete.cases(orders))/nrow(orders)
# 81% missing, but mainly due to postal code which maybe won't matter too much for analysis, maybe MNAR if not in america???
md.pattern(orders)
aggr(orders)



```

## Problem 2: Inventory Management

Retailers that depend on seasonal shoppers have a particularly challenging job when it comes to inventory management. Your manager is making plans for next year's inventory.

He wants you to answer the following questions:

  1. Is there any seasonal sales trend in your company? 
  2. Is there any seasonal trend of **different categories ** of products?

**Note:** Each order has a column called Quantity.

```{r prob2}

colMeans(orders[, c("Quantity", "New_sales", "New_profit")])

colMeans(orders[, c("Quantity", "New_sales", "New_profit")])


time <- orders %>%
  group_by(Order.Date) %>%
  summarise(daily_quantities=sum(Quantity), 
            daily_sales=sum(New_sales), 
            daily_profit=sum(New_profit))

ggplot(time, aes(Order.Date, daily_quantities)) + 
  geom_line() +
  xlab("Time") + 
  ylab("Daily Orders Quantity") +
  theme_bw() +
  geom_smooth()





#histogram of dates
ggplot(orders, aes(x = Order.Date)) + geom_histogram()
#seems to be spikes in orders at end of year for holidays
ggplot(orders, aes(x = Order.Date, y = Quantity)) + geom_bar(stat='identity')
#pattern in quantity of order seems harder to detect. 
#possibly lower quantities in first quarter but then goes up after that


#add a month column to examine data by month
orders$Month = format(orders$Order.Date, '%m')
sort(unique(orders$Month)) #It worked!
class(orders$Month)

ggplot(orders, aes(x = Month)) + geom_bar()
#peaks in june, sept, nov, dec
#low jan - apr & july

#by categories
# orders_bycat = orders %>% 
#   group_by(Category)
# line plot of three categories shows how trends follow each other 
ggplot(orders, aes(x = Month)) + geom_bar(aes(fill = Category))
ggplot(orders, aes(x = Month)) + geom_bar(aes(fill = Category), position = 'dodge')
#office supplies are hottest item but all categories follow same trend
ggplot(orders, aes(x = Month)) + geom_bar(aes(fill = Sub.Category), position = "fill")
#subcategory also pretty consistent throughout
```

## Problem 3: Why did customers make returns?
Your manager required you to give a brief report (**Plots + Interpretations**) on returned orders from the **Returns** dataset.

1. How much profit did we lose for each year?
2. How many customer returned more than once? more than 10 times? 
3. Which regions are more likely to return orders?
4. Which categories (sub-categories) of products are more likely to be returned?

*Hint*: 

1. Import **Returns.csv**
2. Merge the **Returns** dataframe you imported with the **Orders** dataframe.

```{r prob3}
# Fill in your code here
returns = fread('./data/Returns.csv')
head(returns)
names(returns)
names(orders)
#rename order id col
colnames(returns)[2] = "Order.ID"

# merge two data frame on order.ID (inner)
merged_df_inner = merge(orders, returns, by = "Order.ID")
head(merged_df_inner)
nrow(merged_df_inner)

#Q1: How much profit did we lose per year?
merged_df_inner$Year =  format(merged_df_inner$Order.Date, '%y')
merged_df_inner %>% group_by(Year) %>%  summarise(Profit = sum(Profit, na.rm = TRUE))
#between 10000 and 20000 per year


#????why does mine disagree with his

#although I saw some negative profits in there so I'm not sure how to handle those ??????????????????????????


#Q2: How many customers returned more than once
customer_count = merged_df_inner %>% 
  group_by(Customer.ID) %>% 
  count()
customer_count = arrange(customer_count, desc(n))
nrow(filter(customer_count, n> 1))   #11350 out of 51000 (he only got 1159 customers.....weird) total orders??? wow!
nrow(filter(customer_count, n> 10))   #only 260 (still seems extreme)

#Q3: which regions more likely to return orders
returns[!is.na(returns$Region) & returns$Region == 'Eastern Canada','Region'] <- 'Canada'
returns[!is.na(returns$Region) & returns$Region == 'Western Canada','Region'] <- 'Canada'

merged_df_outer = merge(orders, returns, by = "Order.ID", all = TRUE)  #outer join now
regions = merged_df_outer %>% group_by(Region.y) %>% count()

regions_total = merged_df_outer %>% group_by(Region.x) %>% count()
merged_regions = merge(regions, regions_total, by.x = "Region.y", by.y = "Region.x", all = TRUE)
#canada vs. Eastern and western Canada # go back and rename and combine

merged_regions = merged_regions %>% mutate(percent = round(n.x / n.y * 100,1))
merged_regions %>% arrange(desc(percent))
#Western US, Eastern Asia, Southern Europe, Southern Africa, Southern US all above 5% returns


#Q4: categories/subcategories more likely to be returned 

categories =  merged_df_inner%>% group_by(Category) %>% count()
categories_total = merged_df_outer %>%  group_by(Category) %>%  count()

merged_cats = merge(categories, categories_total, by = "Category", all = TRUE)
merged_cats = merged_cats %>% mutate(percent = round(n.x / n.y * 100,1))
merged_cats %>% arrange(desc(percent))
#all categories show 4.3/4.4% returns


subcategories =  merged_df_inner%>% group_by(Sub.Category) %>% count()
subcategories_total = merged_df_outer %>%  group_by(Sub.Category) %>%  count()

merged_subcats = merge(subcategories, subcategories_total, by = "Sub.Category", all = TRUE)
merged_subcats = merged_subcats %>% mutate(percent = round(n.x / n.y * 100,1))
merged_subcats %>% arrange(desc(percent))

#Labels and tables most returned
#fasteers and appliances least returned

```